Compositional equivalences based on Open pNets

### 2020-07-14 2016-Weak-FH-Bisim/WeakBisim.tex
related work

\paragraph{Open and Compositional systems}

\TODO{retravailler le paragraphe ci dessous: faire ref a Facs-2013 (reconfig ds pNets) - virer 14 (probabilistic) ?}
In
\cite{Johnson:2013:CBSE,Johnson:2014:qosa}, the authors investigate several methodologies for the compositional verification of software systems, with the aim to verify reconfigurable component systems. To improve scaling and compositionality, the authors  decouple the verification problem that is to be resolved by a SMT (satisfiability modulo theory) solver into independent sub-problems on independent sets of variables. These works clearly highlight the interest of incremental and compositional verification in a very general setting. By adding more structure to the composition model and designing a dedicated bisimulation theory, we show here how to enforce a compositional proof system that is more powerful than independent sets of variables. Our theory has also been encoded into an SMT solver and it would be interesting to investigate how the examples of evolving systems studied by the authors could be encoded into pNet and verified by our framework. However, the models of Johnson et al. are quite different from ours, in particular they are much less structured, and translating them is clearly outside the scope of this article. Additionally, the more compositional constraint resolution provided by the authors could also be interesting to make our SMT proofs scale better (inside a single pNet for example), the integration of the Z3 framework described in \cite{Johnson:2014:qosa} with our proof theory for open pNets would also be worth investigating from a practical point of view. One major different however is that pNets in their current state are not much adapted to probabilistic verification contrarily to \cite{Johnson:2014:qosa} but they are better suited to verify complex behavioural properties.

@@

\paragraph{Symbolic and data-sensitive systems}

Lin present a symbolic bisimulation between symbolic transition graph with assignments (STGA), as mentioned in \emph{Introduction}, bring us lots of inspirations.

@@

We are currently extending this work,  looking at  both further properties of 
FH-bisimulation, but also
the relations with existing equivalences on closed systems.
We also plan to apply open pNets to the study of complex composition
operators in a symbolic way, for example in the area of parallel
skeletons, or distributed algorithms.
We have developed tool support for computing the
symbolic semantics in term of open-automata \cite{QBMZ-AVOCS18}, and have
developed algorithms to check strong FH-bisimulation \cite{hou:hal-02406098}.  
Naturally we are now working toward an implementation of weak FH-Bisimulation. 
The challenges here, in the context of uour symbolic systems, is not so much algorithmic complexity,
as was the case with classical weak bisimulation on finite models, but decidability and termination.
Two main directions should be explored, either with an explicite construction of the weak transition, but this step in
itself may introduce non-termination, or a direct implementation of the weak bisimulation definition, without
constructing the weak automaton. We can expect that optimisations here in the symbolic setting would be quite different from the classical case.

@@

The three for all statements (on $J_a$, $J_b$, and $J_c$) can be concatenated using $\dotcup$, the list union lifted to indexed sets (if $\gamma=\gamma'$ and $\gamma''=\gamma'''$ then $\gamma\dotcup\gamma''=\gamma'\dotcup\gamma'''$).

>>>>

\subsubsection*{Open and Compositional systems}
\TODO{retravailler le paragraphe ci dessous: faire ref a Facs-2013 (reconfig ds pNets) - virer 14 (probabilistic) ?} \cite{GasparHM:2013}
In
\cite{Johnson:2013:CBSE,Johnson:2014:qosa}, the authors investigate several methodologies for the compositional verification of software systems, with the aim to verify reconfigurable component systems. To improve scaling and compositionality, the authors  decouple the verification problem that is to be resolved by a SMT (satisfiability modulo theory) solver into independent sub-problems on independent sets of variables. These works clearly highlight the interest of incremental and compositional verification in a very general setting. By adding more structure to the composition model and designing a dedicated bisimulation theory, we show here how to enforce a compositional proof system that is more powerful than independent sets of variables. Our theory has also been encoded into an SMT solver and it would be interesting to investigate how the examples of evolving systems studied by the authors could be encoded into pNet and verified by our framework. However, the models of Johnson et al. are quite different from ours, in particular they are much less structured, and translating them is clearly outside the scope of this article. Additionally, the more compositional constraint resolution provided by the authors could also be interesting to make our SMT proofs scale better (inside a single pNet for example), the integration of the Z3 framework described in \cite{Johnson:2014:qosa} with our proof theory for open pNets would also be worth investigating from a practical point of view. One major different however is that pNets in their current state are not much adapted to probabilistic verification contrarily to \cite{Johnson:2014:qosa} but they are better suited to verify complex behavioural properties.

@@

\subsubsection*{Symbolic and data-sensitive systems}
In \cite{Abriola:2018} the authors present the notion of ''data-aware'' bisimulations on data graphs, computation of such bisimulations is 
studied based on XPath logical language extended with tests for data equality. As mentioned in the \emph{Introduction}, the work that brought us a lot of inspirations are those of Lin et al. \cite{HENNESSY:95, Lin:96}. They developed the theory of symbolic transition graphs (STG), and the associated symbolic (early and late) bisimulations, they also study STGs with assignments which can be a model for message-passing processes.

%Lin present a symbolic bisimulation between symbolic transition graph with assignments (STGA), as mentioned in \emph{Introduction}, bring us lots of inspirations.

@@

We are currently extending this work,  looking at  both further properties of 
FH-bisimulation, but also
the relations with existing equivalences on closed systems.
We also plan to apply open pNets to the study of complex composition
operators in a symbolic way, for example in the area of parallel
skeletons, or distributed algorithms.
We have developed tool support for computing the
symbolic semantics in term of open-automata \cite{QBMZ-AVOCS18}, and have
developed algorithms to check strong FH-bisimulation \cite{hou:hal-02406098}.  
Naturally we are now working toward an implementation of weak FH-Bisimulation. 
The challenges here, in the context of our symbolic systems, is not so much algorithmic complexity,
as was the case with classical weak bisimulation on finite models, but decidability and termination.
Two main directions should be explored, either with an explicit construction of the weak transition, but this step in
itself may introduce non-termination, or a direct implementation of the weak bisimulation definition, without
constructing the weak automaton. We can expect that optimisations here in the symbolic setting would be quite different from the classical case.

@@

The three for all statements (on $J_a$, $J_b$ and $J_c$) can be concatenated using $\dotcup$, the list union lifted to indexed sets (if $\gamma=\gamma'$ and $\gamma''=\gamma'''$ then $\gamma\dotcup\gamma''=\gamma'\dotcup\gamma'''$).
